The Power of Images at Aphrodisias: How Digital Resources Can Transform our assessment of Palaeography

### 2016-02-09 source/10/Paper proposal for EAGLE final.tex
graham and licciardello biblio updated
lower case titles
graham updated

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{FACE VALUE: THE POWER OF IMAGES AT APHRODISIAS. How Digital Resources Can Transform our use of Palaeography in understanding Inscriptions.}
\author[war]{Abigail S. Graham\corref{first}}
\address[war]{The University of Warwick}
\cortext[first]{Corresponding author. Email: Abigail.graham@warwick.ac.uk}
\date{2015-11-16}
\begin{document}
 \maketitle
\begin{abstract}
This paper will consider how the publication of a large digital corpus The Inscriptions of Aphrodisias \citet{ReynoldsRouecheBodard2007} has shaped the assessment of inscriptions, particularly regarding palaeography and the dating
of inscriptions. A case study of dedicatory inscriptions from the Temple of Aphrodite at Aphrodisias will explore how
our approach to palaeography and dating has evolved with digital resources, identifying areas where challenges remain
and considering how improvements could be made in both our approach to and in publication of epigraphic materials.

@@

\noindent  As the co-ordinator and lecturer on three graduate level courses of Roman Epigraphy, I am invariably asked the same
question: how do you use letterforms to date inscriptions? My answer is always the same: ``very carefully''. Studying
palaeography within the discipline of ancient epigraphy can be a journey into thorny hedge where one can easily ``fall
into that category of human endeavor known as stylistic attribution and inevitably involve subjectivity''.\footnote{\citet[3-4]{tracy1995}} Stephen Tracy advises that palaeographic surveys should be carried out with caution: ``Caveat Lector
must needs be our motto''. The study of palaeography is problematic on a number of levels, both in the way it is
defined and the ways in which it is employed. Many scholars dismiss lettering as a means of dating and they make an
important point: dating by a single criterion, especially a stylistic one, is somewhat precarious. When studied in
isolation, letterforms present stylistic variations that may be characteristic of a specific individual, workshop or an
urban area. While observations about carving techniques can be helpful in specific case studies, they are more
problematic when applied generally on a broader scale (e.g. to larger geographical areas or time trame) where
archaisms, local styles, and variations can create distortions.\footnote{\citet{susini1973}; \citet{petrucci1993}; \citet[163-181]{Manzella1995}; \citet[433]{cooley2012}; \citet[122-125]{bruun2015}.} Similar caution would be applied in dating a
sculpture on the basis of a hair fragment. Analysis of statues considers a number of factors, material, hairstyle,
drapery and/or context. Inscriptions are doubly difficult, as they fall into categories of both text and an object.
While there is potentially more information, there is also a greater chance that it will be contradictory. Thus the use
of lettering, an ``imprecise science'', is better used in combination with a number of different factors.\footnote{\citet[26-9]{harris1989}; \citet[3-5]{bodel2001}; \citet[393-418]{Manzella2007} and \citet[432-435]{cooley2012}.} Despite the
aforementioned limitations, a number of informative palaeographic studies have been produced. The success of these is
based on detailed commentaries on a specific corpus of material, a transparent methodology and the incorporation of
numerous high quality images.\footnote{ For example: \citet{Gordon1957} offers more than 50 plates and figures, \citet{Manzella1987} provides 218 illustrations, \citet{tracy1990,tracy1995,tracy2003} uses over 60 plates and figures in each volume.}

@@

The advent of online corpora have increased both the access and the development of discussions regarding epigraphic
monumentality, including new methodologies, approaches as well as attempts to redefine genre classifications and
terminology.\footnote{\citet[22-39]{woolf1996}, \citet{eck2010}, \citet[1-10]{panciera2012}, \citet[1-17]{graham2013}. For a recent collection and analysis of epigraphic databases, see \citet[78-85]{Elliot2015}.} A
number of recent studies have used changes in the appearance of inscriptions over time, such as the use of different
media, decorative and paratextual elements\footnote{On marks in the text see \citet[26]{susini1973} and \citet[143-155]{cooley2014}.} (e.g. ivy leaves as interpuncts (hederae distinguentes),\footnote{\citet[293-303]{hommel1970}}
abbreviations\footnote{\citet[111]{Gordon1957}} and spacing between words\footnote{On using
space to reconstruct inscriptions see \citet[195-226]{alfoldy1995}, \citet[151-176]{grasby2002}.} alongside lettering, as dating
criteria.\footnote{ For an extensive list of dating conventions cf. \citet[Chapter 20]{Manzella1987}. }
Panciera’s recent article, in particular, advocates the significance of the public context and visibility of
inscriptions.\footnote{ ``I would propose to regard as an ‘inscription’ any particular type of written human
communication of the sort that we would today call unidirectional…not being addressed to a person or to a group but to
a collectivity, and for this reason is made with the location, writing technique, graphic form and impagination, mode
and register of expression chosen because they are most suitable to the attainment of its intended goal.'' \citep[8]{panciera
2012}.} In this vein, it is worth considering how the physical characteristics of an inscription belong within a
broader assessment of a culture of writing. Is palaeography the study of letterforms alone? Was a focus on letterforms
a deliberate choice or a product of the traditional constraints in accessibility and publication of epigraphic
materials? The following case study from Aphrodisias will explore these questions further.

@@

\noindent The focus of this paper is a series of rather unimpressive column dedications from the temple of Aphrodite at
Aphrodisias. While the information recorded in the texts is unremarkable, the journey of these inscriptions from
their publication as part of an epigraphic corpus to a recent digital publication reveals an epic transformation in the
format and approach to these materials. This study will begin with brief overview of previous published editions
(e.g. \citet{mama8}) followed by an assessment of the information available on the current Inscriptions of
Aphrodisias (2007) website. Through careful assessment of text (including formula, vocabulary and spelling) and its
presentation (the arrangement of the text, use of decoration and spaces, as well as lettering) in the images provided,
this survey will demonstrate how the availability of published images and inclusion of dating criteria have increased
the amount of information available whilst also adding clarity to the process of dating an inscription. By examining
how we define and use palaeography to evaluate inscriptions on the Inscriptions of Aphrodisias website \citep{ReynoldsRouecheBodard2007}, we
can observe how the discipline has evolved and what changes may be possible in approaches and the publication of
epigraphic material.

@@

These texts can be dated through a number of different criteria. Contextual association with construction of the
temple, dates between the end of the 1st century BC and the early 1st century AD, though the land was clearly in use
well before this time. \footnote{\citet[41-42]{smith1996} and \citet[37]{reynolds1990}. References to Sulla’s dedications are in
Appian BC 1.97. Caesar and Augustus’ acknowledgements of the sacred space are evident in IAph 8.27, 8.31 and 1.1. }
This is corroborated by coins depicting the temple, which date from 2 BC- 14 AD.\footnote{ Coin type 41 depicts
Augustus (OBV) and the temple of Aphrodite (REV) in MacDonald 1992, Plate V R131. } Prosopography is also informative:
Gaius Julius Zoilos (freed by Caesar or Octavian), who dedicated the theatre at Aphrodisias ca. 28 BC, also dedicated
the naos of the temple (IAph 1.2), perhaps posthumously. \footnote{ The titles on the architrave inscription, soter and
euergetes, are absent from Zoilos’ other inscriptions and may imply a posthumous dedication, perhaps by the boule and
demos \citep[38]{reynolds 1990}.} The benefactor of the columns in this case study, Eumachos Diogenes, also comes from an
established family that flourished well into the 2nd century AD.\footnote{18 The success of Eumachus Diogenes’ family,
which included the first known Aphrodisian to hold a procuratorship in the 2nd century AD, reveals an enduring
significance for his family monuments in the city \citep[327-334]{reynolds1999}.} Finally, the formula of the text the
inscription and the vocabulary, particularly the term philokaisar, suggest a late Republican or Augustan
date.\footnote{ The formula of the dedication, which lists the benefactor first is evident throughout the Hellenistic
period in Aphrodisias (and Asia Minor) until the early Imperial Period, after which a new formula (beginning with
recipients (e.g. Aphrodite and Imperial recipients) is predominant (cf. \citet[4-7]{graham2013}). The Augustan and early
Tiberian uses of the philokaisar are known from a dedication at Ioulis dated ca. 27 BC - AD 14 (SEG XLVIII (1998) no.
1129) and in a monument to Ti. Cl. Drusus in Patara (SEG XLIV (1994) no. 1205). For significance and date of the title
see \citet[101-105]{buraselis2001}.} Before encountering the inscriptions face to face, the dating, function and meaning of
these inscriptions appears to be quite straightforward. So let us examine the experience of viewing these
inscriptions various published formats.

@@

\begin{quotation}
\noindent MAMA 437\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηναγό-\\
ρου τοῦ Ἀθηναγόρου\\
τοῦ Εὐμάχου Διογένη-\\
ς φιλόκαισαρ καὶ Ἀμιὰς\\
Διονυσίου φύσι δὲ Ἀδρά<σ>του\\
τοῦ Μόλωνος Ὀλυνπιὰς\\
τὸν κίονα θεᾷ Ἀφροδίτῃ\\
καὶ τῷ δήμῳ.
\end{otherlanguage*}
\end{quotation}

@@

\subsection{Living in a Digital World: The Publication of Column Dedications in IAph 2007 (IAph 1.4.)}
The Inscriptions of Aphrodisias project was a groundbreaking endeavour both for publishing epigraphic materials and
employing Epi Doc conventions with XML for marking up ancient documents. Its success has inspired a number of
significantly larger projects with international scope and collaboration. The approach to the materials is detailed
with categories covering physical characteristics, history of discovery and bibliography (which is important given the
wealth of superlative scholarship).\footnote{ It is worth noting that websites, however they develop, will remain
supplementary to published scholarship.} In addition to search functions, the page is more interactive, the texts
have been revisited, translations added, as well as commentary with weblinks to parallel texts. The organization of the
inscriptions by context allows the reader to view the inscriptions in context and to gain a better understanding of the
area’s ``epigraphic habit'' (Fig.~\ref{fig:1}).

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img001.png}
\caption{Screenshot of IAph 4 webpage}
\label{fig:1}
\end{figure}

In addition to the format of the webpage, a number of images have been published along with a drawing from Deering’s
notebook from the early 19th century. \ These resources provide an opportunity not only to view the inscription but to
see how it has been studied over time. The dating of the inscription includes a list of the criteria upon which it was
based: lettering, context, prosopography, adding a degree of transparency to the dating process. While the lettering is
not subject to further description, the inclusion of weblinks to parallel texts allows the viewer to compare and
contrast different inscriptions, developing his/her understanding this element.

The published texts of IAph 1.4 and MAMA 437 present one significant variation: the inclusion of space indicators within
the text. This practice allows the reader to see how the use of space relates with the text of the inscription,
particularly in the case of line 4, where challenges in the text (misspelling of Ammias and carrying over of a single
letter from the previous line), can be observed in the inclusion of a vacat at the end of the line.\footnote{ Vacats,
more sparingly employed in late Republican and Augustan inscriptions at Aphrodisias, usually serve a grammatical and/or
decorative function (e.g. giving distinction to a name or key elements/individuals in the text (cf. building
dedications at the Sebasteion IAph 9.1, 9.25, 9.112). This vacat serves little grammatical function, and is likely to
be a result of the arrangement of the text/ omission in Ammias’ name. } The indentation on the last line also reveals
a left orientation of the line which is more common in Late Republican/Augustan texts at Aphrodisias (e.g. IAph 1.1,
1.7, 1.8 and 1.38), as opposed to the ``justified'' approach (indentations on both sides of the text to accentuate a name
or word), which is more common in later Imperial inscriptions at Aphrodisias.

\begin{quotation}
\noindent MAMA 437\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηναγό-\\
ρου τοῦ Ἀθηναγόρου\\
τοῦ Εὐμάχου Διογένη-\\
ς φιλόκαισαρ καὶ Ἀμιὰς\\
Διονυσίου φύσι δὲ Ἀδρά<σ>του\\
τοῦ Μόλωνος Ὀλυνπιὰς\\
τὸν κίονα θεᾷ Ἀφροδίτῃ\\
καὶ τῷ δήμῳ.
\end{otherlanguage*}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img002.jpg}
\caption{Photograph of IAph 1.4}
\label{fig:2}
\end{figure}

Alongside spelling and the arrangement of the inscription, Letterforms are more readily observed. Overall, the lettering
is less uniform than later Imperial dedications, serifs are small and the strokes are both thick and deeply incised.
Letter height varies significantly as does the size of omicrons (though this can be seen in a poorly rendered text
throughout the Roman period). Angular forms such as alpha, lambda, and mu all intersect at the top meeting at sharper
angles than the more square versions of these letters that predominate in Imperial (post Augustan)
inscriptions.\footnote{ Compare with Hadrianic Lettering in the next section of the paper.} \ Letters also bear
stylistic elements of the time: epsilons have a connected middle bar (this is often disconnected in Imperial versions
of this letter), the rhos have small legs, making them appear more like their Latin counterpart, and the omicron has an
oval shape with serifs only at one end of the lower bars.\footnote{ ``Legged'' rhos are observed primarily in Augustan
and Julio-Claudian dedications and also on coins (cf. note 16). This style letterform is increasingly less common after
Flavian period at Aphrodisias. } Comparing these letterforms with parallel dedications at the temple (IAph 1.1, 1.2
(Zoilos’ dedications), 1.7 and 1.8) reveals many similarities. Letterforms alone are not diagnostic but when combined
with other elements, they can add to our understanding of a date.

@@

\begin{quotation}
\noindent IAph 1.5.\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηνα-\\
γόρου τοῦ Ἀθηναγό-\\
ρου τοῦ Εὐμάχου Δι-\\
ογένης Φιλόκαισαρ\\
 καὶ Ἀμμιὰς Διονυσί-\\
ου φύσι δὲ Ἀδράστου\\
τοῦ Μόλωνος Ὀλυν-\\
πιὰς τὸν κίονα θεᾷ\\
Ἀφροδίτῃ καὶ τῷ
\end{otherlanguage*}
vac. \begin{otherlanguage*}{greek}
Δήμῳ
\end{otherlanguage*} vac.
\end{quotation}

An image (Fig.~\ref{fig:3}) illustrates further differences between the appearance of this inscription
and the comparative materials from the site (IAph 1.1, 1.2, 1.4, 1.7 and 1.8).

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img003.jpg}
\caption{Photograph of IAph 1.5}
\label{fig:3}
\end{figure}

The high quality image reveals square and regular letterforms (e.g. omicrons are less oval and more precisely rendered
throughout) with a clear contrast between slender strokes and large deep triangulate serifs. There are no legged rhos,
the middle bars of the epsilons are separated from the stem, and the omegas are circular with two large bars that are
heavily serifed on both ends (e.g. compare the omega on the bottom line of IAph 1.4 with its counterpart of IAph.1.5).
These traits along with others (e.g. letter size and use of spacing noted above) are commonly attributed to late 1st
/2nd c. AD inscriptions at Aphrodisias, and are absent from comparative Augustan materials on the site (e.g. IAph 1.1,
1.2, 1.4, 1.7 and 1.8, including a Julio-Claudian dedication 1.102). Although one must be careful with stylistic
factors, especially when archaisms can be used, it is worth noting when a number of features that are attributed to a
later period, seem to suddenly emerge nearly a century early.\footnote{ Archaising and imitation of earlier lettering
is evident in at least one inscription (IAPH 13.116) \citet[155-166]{reynolds1982}). } Any one of these features, letter
variations, stylistic letterforms, use of serifs, indentations in the text, correction of errors in a previous text,
would not stand well alone, but taken together, a more compelling case can be made for the reassessment of this
inscription’s date.

The website records the date of this inscription just like IAph 1.4: citing ``context, lettering and prosopography.'' For
those trying to gain an understanding of letterforms at Aphrodisias, this is somewhat confusing and it is not the only
example in which the lettering and organization of the inscription do not match date provided.\footnote{ Honours for
Zoilos (IAph 8.203) is dated by ``prosopography'' to the 1st c. BC but the letters are square Imperial forms and do not
match Zoilos’ other dedications (8.1, 1.1,1.2). Honours for P. and M. Vinicius (IAph 3.101) are dated as ``Augustan'' and
``Tiberian'' by ``lettering'' though prosopography is known (there is some controversy cf. Reynolds 1982,175). Stylistic
elements (arrangement, spacing and decoration), particularly in M. Vinicius’ base, reflect qualities of
Claudian-Flavian period inscriptions (which is not excluded by prosopographical dates). In both cases, conflicting
dating criteria are omitted. } The date of the text is not incorrect, insofar as it was originally inscribed at
this time and the prosopography as well as the formula of the text support this. However, one struggles to see how
the lettering and/or the arrangement of the text corroborate this date. An answer may be found in further examination
of the inscription’s context and prosopography. In addition to a number of earthquakes, new temenos of the temple was
added under the emperor Hadrian, a time during which Athenagoras’ descendants were alive and prospering in
Aphrodisias.\footnote{\citep[43]{smith1995}. For an analysis of the temenos plan, see \citet[66-74]{doruk1990}. For Athenagoras descendants, cf. note 18 and \citet[327-334]{reynolds1999b}.}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img004.jpg}
\caption{Photograph of IAph 1.174}
\label{fig:4}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img005.jpg}
  \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img006.jpg}
\caption{Photograph of I.Aph 1.6}
\label{fig:5}
\end{figure}

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{The Power of Images at Aphrodisias: 
How Digital Resources Can Transform our assessment of Palaeography}
\author[war]{Abigail S. Graham\corref{first}}
\address[war]{The University of Warwick}
\cortext[first]{Corresponding author. Email: Abigail.graham@warwick.ac.uk}
\date{2015-11-16}
\begin{document}
 \maketitle
\begin{abstract}
This paper will consider how the publication of a large digital corpus The Inscriptions of Aphrodisias \citet{ReynoldsRouecheBodard2007} has shaped the assessment of inscriptions, particularly regarding palaeography and the dating
of inscriptions. A case study of dedicatory inscriptions from the Temple of Aphrodite at Aphrodisias will explore how
our approach to palaeography and dating has evolved with digital resources, identifying areas where challenges remain
and considering how improvements could be made in both our approach to and in publication of epigraphic materials.

@@

\noindent  As the co-ordinator and lecturer on three graduate level courses of Roman Epigraphy, I am invariably asked the same
question: how do you use letterforms to date inscriptions? My answer is always the same: ``very carefully''. Studying
palaeography within the discipline of ancient epigraphy can be a journey into thorny hedge where one can easily ``fall
into that category of human endeavor known as stylistic attribution and inevitably involve subjectivity''.\footnote{\citet[3-4]{tracy1995}} Stephen Tracy advises that palaeographic surveys should be carried out with caution: ``Caveat Lector
must needs be our motto''. The study of palaeography is problematic, as demonstrated by W. Eck and C. Roueché during the conference, on a number of levels, both in the way it is
defined and the ways in which it is employed. Many scholars dismiss lettering as a means of dating and they make an
important point: dating by a single criterion, especially a stylistic one, is somewhat precarious. When studied in
isolation, letterforms present stylistic variations that may be characteristic of a specific individual, workshop or an
urban area. While observations about carving techniques can be helpful in specific case studies, they are more
problematic when applied generally on a broader scale (e.g. to larger geographical areas or time trame) where
archaisms, local styles, and variations can create distortions.\footnote{\citet{susini1973}; \citet{petrucci1993}; \citet[163-181]{Manzella1995}; \citet[433]{cooley2012}; \citet[122-125]{bruun2015}.} Similar caution would be applied in dating a
sculpture on the basis of a hair fragment. Analysis of statues considers a number of factors, material, hairstyle,
drapery and/or context. Inscriptions are doubly difficult, as they fall into categories of both text and an object.
While there is potentially more information, there is also a greater chance that it will be contradictory. Thus the use
of lettering, an ``imprecise science'', is better used in combination with a number of different factors.\footnote{\citet[26-9]{harris1989}; \citet[3-5]{bodel2001}; \citet[393-418]{Manzella2007} and \citet[432-435]{cooley2012}.} Despite the
aforementioned limitations, a number of informative palaeographic studies have been produced. The success of these is
based on detailed commentaries on a specific corpus of material, a transparent methodology and the incorporation of
numerous high quality images.\footnote{ For example: \citet{Gordon1957} offers more than 50 plates and figures, \citet{Manzella1987} provides 218 illustrations, \citet{tracy1990,tracy1995,tracy2003} uses over 60 plates and figures in each volume.}

@@

The advent of online corpora have increased both the access and the development of discussions regarding epigraphic
monumentality, including new methodologies, approaches as well as attempts to redefine genre classifications and
terminology.\footnote{\citet[22-39]{woolf1996}, \citet{eck2010}, \citet[1-10]{panciera2012}, \citet[1-17]{graham2013}. For a recent collection and analysis of epigraphic databases, see \citet[78-85]{Elliot2015}.} A
number of recent studies have used changes in the appearance of inscriptions over time, such as the use of different
media, decorative and paratextual elements\footnote{On marks in the text see \citet[26]{susini1973} and \citet[143-155]{cooley2014}.} (e.g. ivy leaves as interpuncts (hederae distinguentes),\footnote{\citet[293-303]{hommel1970}}
abbreviations\footnote{\citet[111]{Gordon1957}} and spacing between words\footnote{On using
space to reconstruct inscriptions see \citet[195-226]{alfoldy1995}, \citet[151-176]{grasby2002}.} alongside lettering, as dating
criteria.\footnote{ For an extensive list of dating conventions cf. \citet[Chapter 20]{Manzella1987}. }
Panciera’s recent article, in particular, advocates the significance of the public context and visibility of
inscriptions.\footnote{ ``I would propose to regard as an ‘inscription’ any particular type of written human
communication of the sort that we would today call unidirectional…not being addressed to a person or to a group but to
a collectivity, and for this reason is made with the location, writing technique, graphic form and impagination, mode
and register of expression chosen because they are most suitable to the attainment of its intended goal.'' \citep[8]{panciera2012}.} In this vein, it is worth considering how the physical characteristics of an inscription belong within a
broader assessment of a culture of writing. Is palaeography the study of letterforms alone? Was a focus on letterforms
a deliberate choice or a product of the traditional constraints in accessibility and publication of epigraphic
materials? The following case study from Aphrodisias will explore these questions further.

@@

\noindent The focus of this paper is a series of rather unimpressive column dedications from the temple of Aphrodite at
Aphrodisias. While the information recorded in the texts is unremarkable, the journey of these inscriptions from a paper epigraphic corpus to their recent publication as a digital corpus reveals an epic transformation in the
format and approach to these materials. This study will begin with brief overview of previous published editions
(e.g. \citet{calder1962}) followed by an assessment of the information available on the current Inscriptions of
Aphrodisias (2007) website. Through careful assessment of text (including formula, vocabulary and spelling) and its
presentation (the arrangement of the text, use of decoration and spaces, as well as lettering) in the images provided,
this survey will demonstrate how the availability of published images and inclusion of dating criteria have increased
the amount of information available whilst also adding clarity to the process of dating an inscription. By examining
how we define and use palaeography to evaluate inscriptions on the Inscriptions of Aphrodisias website \citep{ReynoldsRouecheBodard2007}, we
can observe how the discipline has evolved and what changes may be possible in approaches and the publication of
epigraphic material.

@@

These texts can be dated through a number of different criteria. Contextual association with construction of the
temple, dates between the end of the 1st century BC and the early 1st century AD, though the land was clearly in use
well before this time. \footnote{\citet[41-42]{smith1996} and \citet[37]{reynolds1990}. References to Sulla’s dedications are in
Appian BC 1.97. Caesar and Augustus’ acknowledgements of the sacred space are evident in IAph 8.27, 8.31 and 1.1. }
This is corroborated by coins depicting the temple, which date from 2 BC- 14 AD.\footnote{ Coin type 41 depicts
Augustus (OBV) and the temple of Aphrodite (REV) in MacDonald 1992, Plate V R131. } Prosopography is also informative:
Gaius Julius Zoilos (freed by Caesar or Octavian), who dedicated the theatre at Aphrodisias ca. 28 BC, also dedicated
the naos of the temple (IAph 1.2), perhaps posthumously. \footnote{ The titles on the architrave inscription, soter and
euergetes, are absent from Zoilos’ other inscriptions and may imply a posthumous dedication, perhaps by the boule and
demos \citep[38]{reynolds1990}.} The benefactor of the columns in this case study, Eumachos Diogenes, also comes from an
established family that flourished well into the 2nd century AD.\footnote{The success of Eumachus Diogenes’ family,
which included the first known Aphrodisian to hold a procuratorship in the 2nd century AD, reveals an enduring
significance for his family monuments in the city \citep[327-334]{reynolds1999}.} Finally, the formula of the text the
inscription and the vocabulary, particularly the term philokaisar, suggest a late Republican or Augustan
date.\footnote{ The formula of the dedication, which lists the benefactor first is evident throughout the Hellenistic
period in Aphrodisias (and Asia Minor) until the early Imperial Period, after which a new formula (beginning with
recipients (e.g. Aphrodite and Imperial recipients) is predominant (cf. \citet[4-7]{graham2013}). The Augustan and early
Tiberian uses of the philokaisar are known from a dedication at Ioulis dated ca. 27 BC - AD 14 (SEG XLVIII (1998) no.
1129) and in a monument to Ti. Cl. Drusus in Patara (SEG XLIV (1994) no. 1205). For significance and date of the title
see \citet[101-105]{buraselis2000}.} Before encountering the inscriptions face to face, the dating, function and meaning of
these inscriptions appears to be quite straightforward. So let us examine the experience of viewing these
inscriptions various published formats.

@@

\begin{quotation}
\noindent MAMA 437\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηναγό-\\
ρου τοῦ Ἀθηναγόρου\\
τοῦ Εὐμάχου Διογένη-\\
ς φιλόκαισαρ καὶ Ἀμιὰς\\
Διονυσίου φύσι δὲ Ἀδρά\textlatin{<}σ\textlatin{>}του\\
τοῦ Μόλωνος Ὀλυνπιὰς\\
τὸν κίονα θεᾷ Ἀφροδίτῃ\\
καὶ τῷ δήμῳ.
\end{otherlanguage*}
\end{quotation}

@@

\subsection{Living in a Digital World: The Publication of Column Dedications in IAph 2007 (IAph 1.4.)}
The Inscriptions of Aphrodisias project was a groundbreaking endeavour both for publishing epigraphic materials and
employing Epi Doc conventions with XML for marking up the text of an inscription. Its success has inspired a number of
significantly larger projects with international scope and collaboration. The approach to the materials is detailed
with categories covering physical characteristics, history of discovery and bibliography (which is important given the
wealth of superlative scholarship).\footnote{ It is worth noting that websites, however they develop, will remain
supplementary to published scholarship such as J. Reynolds wonderfully detailed \emph{Aphrodisias and Rome} (1982).} In addition to search functions, the page is more interactive, the texts
have been revisited, translations added, as well as commentary with weblinks to parallel texts. The organization of the
inscriptions by context allows the reader to view the inscriptions in context and to gain a better understanding of the
area’s ``epigraphic habit'' (Fig.~\ref{fig:1}).

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img001.png}
\caption{Screenshot of IAph 4 webpage}
\label{fig:1}
\end{figure}

\newpage
In addition to the format of the webpage, a number of images have been published along with a drawing from Deering’s
notebook from the early 19th century. \ These resources provide an opportunity not only to view the inscription but to
see how it has been studied over time. The dating of the inscription includes a list of the criteria upon which it was
based: lettering, context, prosopography, adding a degree of transparency to the dating process. While the lettering is
not subject to further description, the inclusion of weblinks to parallel texts allows the viewer to compare and
contrast different inscriptions, developing his/her understanding this element.

The published texts of IAph 1.4 and MAMA 437 present one significant variation: the inclusion of space indicators within
the text. This practice allows the reader to see how the use of space relates with the text of the inscription,
particularly in the case of line 4, where challenges in the text (misspelling of Ammias and carrying over of a single
letter from the previous line), can be observed in the inclusion of a vacat at the end of the line.\footnote{ Vacats,
more sparingly employed in late Republican and Augustan inscriptions at Aphrodisias, usually serve a grammatical and/or
decorative function (e.g. giving distinction to a name or key elements/individuals in the text (cf. building
dedications at the Sebasteion IAph 9.1, 9.25, 9.112)). This vacat serves little grammatical function, and is likely to
be a result of the arrangement of the text/ omission in Ammias’ name. } The indentation on the last line also reveals
a left orientation of the line which is more common in Late Republican/Augustan texts at Aphrodisias (e.g. IAph 1.1,
1.7, 1.8 and 1.38), as opposed to the ``justified'' approach (indentations on both sides of the text to accentuate a name
or word), which is more common in later Imperial inscriptions at Aphrodisias. 
\newpage
\begin{quotation}
\noindent MAMA 437\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηναγό-\\
ρου τοῦ Ἀθηναγόρου\\
τοῦ Εὐμάχου Διογένη-\\
ς φιλόκαισαρ καὶ Ἀμιὰς\\
Διονυσίου φύσι δὲ Ἀδρά\textlatin{<}σ\textlatin{>}του\\
τοῦ Μόλωνος Ὀλυνπιὰς\\
τὸν κίονα θεᾷ Ἀφροδίτῃ\\
καὶ τῷ δήμῳ.
\end{otherlanguage*}

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img002.jpg}
\caption{Photograph of IAph 1.4}
\label{fig:2}
\end{figure}

Alongside spelling and the arrangement of the inscription, Letterforms are more readily observed. Overall, the lettering
is less uniform than later Imperial dedications, serifs are small and the strokes are both thick and deeply incised.
Letter height varies significantly as does the size of omicrons (though this can be seen in a poorly rendered text
throughout the Roman period). Angular forms such as alpha, lambda, and mu all intersect at the top meeting at sharper
angles than the more square versions of these letters that predominate in Imperial (post Augustan)
inscriptions.\footnote{ Compare with Hadrianic Lettering in the next section of the paper.} \ Letters also bear
stylistic elements of the time: epsilons have a connected middle bar (this is often disconnected in Imperial versions
of this letter), the rhos have small legs, making them appear more like their Latin counterpart, and the omicron has an
oval shape with serifs only at one end of the lower bars.\footnote{ ``Legged'' rhos are observed primarily in Augustan
and Julio-Claudian dedications and also on coins (cf. note 15). This style letterform is increasingly less common after
Flavian period at Aphrodisias. } Comparing these letterforms with parallel dedications at the temple (IAph 1.1, 1.2
(Zoilos’ dedications), 1.7 and 1.8) reveals many similarities. Letterforms alone are not diagnostic but when combined
with other elements, they can add to our understanding of a date.

@@

\begin{quotation}
\noindent IAph 1.5.\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηνα-\\
γόρου τοῦ Ἀθηναγό-\\
ρου τοῦ Εὐμάχου Δι-\\
ογένης Φιλόκαισαρ\\
 καὶ Ἀμμιὰς Διονυσί-\\
ου φύσι δὲ Ἀδράστου\\
τοῦ Μόλωνος Ὀλυν-\\
πιὰς τὸν κίονα θεᾷ\\
Ἀφροδίτῃ καὶ τῷ
\end{otherlanguage*} \\
vac. \begin{otherlanguage*}{greek}
Δήμῳ
\end{otherlanguage*} vac.
\end{quotation}

A high quality image (Fig.~\ref{fig:3}) illustrates further differences between the appearance of this inscription
and the comparative materials from the site (IAph 1.1, 1.2, 1.4, 1.7 and 1.8), revealing regular letterforms (e.g. omicrons are less oval and more precisely rendered
throughout) with a clear contrast between slender strokes and large deep triangulate serifs. There are no legged rhos,
the middle bars of the epsilons are separated from the stem, and the omegas are circular with two large bars that are
heavily serifed on both ends (e.g. compare the omega on the bottom line of IAph 1.4 with its counterpart of IAph.1.5).

\newpage

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img003.jpg}
\caption{Photograph of IAph 1.5}
\label{fig:3}
\end{figure}

These traits along with others (e.g. letter size and use of spacing noted above) are commonly attributed to late 1st
/2nd c. AD inscriptions at Aphrodisias, and are absent from comparative Augustan materials on the site (e.g. IAph 1.1,
1.2, 1.4, 1.7 and 1.8, including a Julio-Claudian dedication 1.102). Although one must be careful with stylistic
factors, especially when archaisms can be used, it is worth noting when a number of features that are attributed to a
later period, seem to suddenly emerge nearly a century early.\footnote{ Archaising and imitation of earlier lettering
is evident in at least one inscription (IAPH 13.116) \citet[155-166]{reynolds1982}). } Any one of these features, letter
variations, stylistic letterforms, use of serifs, indentations in the text, correction of errors in a previous text,
would not stand well alone, but taken together, a more compelling case can be made for the reassessment of this
inscription’s date.

The website records the date of this inscription just like IAph 1.4: citing ``context, lettering and prosopography.'' For
those trying to gain an understanding of letterforms at Aphrodisias, this is somewhat confusing and it is not the only
example in which the lettering and organization of the inscription do not match date provided.\footnote{ Honours for
Zoilos (IAph 8.203) is dated by ``prosopography'' to the 1st c. BC but the letters are square Imperial forms and do not
match Zoilos’ other dedications (8.1, 1.1,1.2). Honours for P. and M. Vinicius (IAph 3.101) are dated as ``Augustan'' and
``Tiberian'' by ``lettering'' though prosopography is known (there is some controversy cf. Reynolds 1982,175). Stylistic
elements (arrangement, spacing and decoration), particularly in M. Vinicius’ base, reflect qualities of
Claudian-Flavian period inscriptions (which is not excluded by prosopographical dates). In both cases, conflicting
dating criteria are omitted. } The date of the text is not incorrect, insofar as it was originally inscribed at
this time and the prosopography as well as the formula of the text support this. However, one struggles to see how
the lettering and/or the arrangement of the text corroborate this date. An answer may be found in further examination
of the inscription’s context and prosopography. In addition to a number of earthquakes, new temenos of the temple was
added under the emperor Hadrian, a time during which Athenagoras’ descendants were alive and prospering in
Aphrodisias.\footnote{\citep[43]{smith1995}. For an analysis of the temenos plan, see \citet[66-74]{doruk1990}. For Athenagoras descendants, cf. note 17 and \citet[327-334]{reynolds1999}.}

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img004.jpg}
\caption{Photograph of IAph 1.174}
\label{fig:4}
\end{figure}

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img005.jpg}
  \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img006.jpg}
\caption{Photograph of I.Aph 1.6}
\label{fig:5}
\end{figure}
